---
title: "Assignment 2"
author: "Cameron Atkins"
date: "2023-09-29"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Problem 1

Build a function that takes a vector of size 500 (call the function and argument whatever you would
like). If the vector is normally distributed, find the 95% confidence interval and print that to the console
(use the cat() function). If the vector is not normally distrubted, print a message back to the console
that the vector is not normally distributed. Regardless of distribution, plot the density of the vector.


Run a test case on both functions where the data is normally distributed and one where the
data is not. The best way would be to use some r<distribution>() functions (R has numerous
distributions you can use to create vectors from with random values).
```{r cars}

confInterval <- function(v1){
  vectorMean <- mean(v1)
  standardError <- (qnorm(0.975)*sd(v1))/(sqrt(length(v1)))
  confidence_interval <- c(vectorMean - standardError, vectorMean + standardError)
  cat("95% Confidence Interval:", confidence_interval, "\n")
}

checkIfNormal <- function(v2){
  isNormal <- shapiro.test(v2)

  if (isNormal$p.value >= 0.05){
    cat("Vector is normally distributed\n")
    confInterval(v2)
  }
  
  else{
    cat("Vector is not normally distributed\n")
  }

# The Density Plot of vector passed to function
  plot(density(v2))
}

# Test functions with data that is normal
set.seed(123)  
isNormal <- rnorm(500)
checkIfNormal(isNormal)

# Test functions with data that is not normal
set.seed(456) 
notNormal <- rexp(500, rate = 1)
checkIfNormal(notNormal)
```

## Problem 2
Create a function that accepts a dataset. For each column in the dataset, print a statement “<column
name> is of type: <variable type>”. Test out your function on the dataset attached to this assignment
(show the steps of reading in the dataset, you may need to use the str() function and make a class
conversion for one of the variables).

### Solution
```{r}

# Read in the dataset
SampleData <- read.csv('dfex.csv')

# Structure of Sample data before converting Date Column
str(SampleData)

# Converting Column x6 from chr to date
SampleData$x6 <- as.Date(SampleData$x6)

ColumnNameType <- function(ds){
  columns <- colnames(ds)
  
  for (name in columns){
    type <- class(ds[[name]])
    
    cat(name, "is of type: ", type, "\n")
  }
}

# Structure of Sample data after converting Date Column to type Date
str(SampleData)

# Test the function with the sample dataset
ColumnNameType(SampleData)
```
## Problem 3
What does each function do? What does each function return? Come up with better names for these functions

### Solution
A. function f1 accepts two parameters, a string and prefix. The function returns true if "string" begins
with "prefix" and false if it doesn't. A better name for this function would be beginsWithPrefix. Below is 
an example of fuction f1 in use.
```{r}
f1 <- function(string, prefix) {
  substr(string, 1, nchar(prefix)) == prefix
}

f1("mathematics", "math")
f1("mathematics", "biol")
```
B. function f2 accepts one parameter, a vector or list. If the list has only one element, the function
return NULL. If the list or vector has more than one element, the last element is removed. 

A better name for this function would be removeLastElement.

Below are two instances of function f2 executing.
```{r}
f2 <- function(x) {
  if (length(x) <= 1) return(NULL)
  x[-length(x)]
}
  
names <- c("Djokovic", "Federer", "Nadal", "Tiafoe", "Alcaraz", "Murray")
nums <- c(4005)

f2(names)
f2(nums)


```
C. function f3 accepts two vectors as parameters. Using the built-in replicate function(rep), 
function f3 returns the second vector, y, matching the number of elements in the first vector, x.
If the length of the y is less than the length of x, then elements in y will be repeated to until its length
matches the length of x. If the length of x is greater than the length of y, elements in y will be removed to match the length of x. 

A better name for this function would be matchesLength. 

Below are four instances of f3 executing with vectors of various lengths.
```{r}
f3 <- function(x, y) {
rep(y, length.out = length(x))
}

a <- c(1, 2, 3, 4, 5)
b <- c(6)
c <- c(7, 8, 9, 10, 11, 12, 13, 14, 15)

f3(a, b)
f3(a, c)
f3(b, c)
f3(c,a)
```

